---
title: "CCA integration of all datasets (std Normalization)"
author: "jos"
date: "05/31/2020"
output:
  html_document:
    fig_height: 12
    fig_width: 17
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
# to knit document, use the following command: rmarkdown::render("C:/Users/Hannah Bishop/Desktop/GitHubCode/Octopolis/Orig_All_StdN_CCA.Rmd") ~30m

library(Seurat)
library(Matrix)
library(ggplot2)
library(sctransform)
library(stringr)
library(cowplot) # used for CCA
library(patchwork) # used for CCA
```

## Read in data
```{r, echo = FALSE, eval = FALSE}
all_list <- readRDS("C:/Users/Hannah Bishop/Desktop/MeaS/all_list.rds")
for (i in 1:length(all_list)){
  #all_list[[i]][["percent.mt"]] <- PercentageFeatureSet(all_list[[i]], pattern = "^mitochondria-") #preferred for for loop over original command, leave out for new GTF because poor annotation
  all_list[[i]] <- NormalizeData(all_list[[i]])
  all_list[[i]] <- FindVariableFeatures(all_list[[i]], selection.method = "vst", nfeatures = 2000)
}

saveRDS(all_list, "C:/Users/Hannah Bishop/Desktop/MeaS/all_list_stdN.rds")
all_list <- readRDS("C:/Users/Hannah Bishop/Desktop/MeaS/all_list_stdN.rds")

## Select features for downstream integration, and run PrepSCTIntegration so that all necessary Pearson residuals have been calculated
#all.features <- SelectIntegrationFeatures(object.list = all_list[2:3], nfeatures = 2000) #paper uses 3000 as default but chose to use 2000 like other dataset integration
#all_list <- PrepSCTIntegration(object.list = all_list[2:3], anchor.features = all.features, verbose = TRUE)

## Identify anchors and integrate datasets. Use the standard workflow, with the exception that normalization.method = 'SCT'
all.anchors <- FindIntegrationAnchors(object.list = all_list,
    anchor.features = 2000, dims = 1:30, verbose = TRUE) # can omit dims but takes a really long time to run
allnorm.integrated <- IntegrateData(anchorset = all.anchors, verbose = TRUE)

saveRDS(allnorm.integrated, "C:/Users/Hannah Bishop/Desktop/MeaS/allStdNCCA.rds")
```

## Read in integrated dataset
```{r, echo = FALSE}
all <- readRDS("C:/Users/Hannah Bishop/Desktop/MeaS/allstdNCCA.rds")
DefaultAssay(all) <- "integrated"
```

## Plot features of the object. 
```{r }
VlnPlot(all, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
#plot1 <- FeatureScatter(all, feature1 = "nCount_RNA", feature2 = "percent.mt")
# visualize the relationship between UMI counts and mitochondrial genes (percent)
plot2 <- FeatureScatter(all, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") 
# visualize the relationship between genes (ID as >1 UMI) and mitochondrial genes
#plot1 + plot2
plot2
```

## Run PCA and generate UMAP
```{r }
all <- ScaleData(all)
all <- RunPCA(all, npcs = 30, verbose = FALSE) 
all <- FindNeighbors(all, reduction = "pca", dims = 1:30) #not sure if "reduction" argument is necessary
all <- FindClusters(all, resolution = 1)
all <- RunUMAP(all, reduction = "pca", dims = 1:30)
DimPlot(all, reduction = "umap", label = TRUE)
DimPlot(all, reduction = "umap", group.by = "orig.ident")
```

## Examine cell-cluster information
```{r }
table(Idents(all))
table(all$orig.ident)
prop.table(table(Idents(all)))
table(Idents(all), all$orig.ident)
```

## Build a hierarchical tree
```{r }
all.tree <- BuildClusterTree(all, reorder = TRUE, reorder.numeric = TRUE, slot = "scale.data", verbose = TRUE, dims = 1:30)
PlotClusterTree(all.tree, label = TRUE)
```

```{r }
all.tree <- RunUMAP(all.tree, dims = 1:30)
DimPlot(all.tree, reduction = "umap", label = TRUE)
table(Idents(all.tree)) 
table(all.tree$orig.ident) 
prop.table(table(Idents(all.tree))) 
table(Idents(all.tree), all.tree$orig.ident) 
```

## Plot features
````{r, echo = FALSE}
yfg <- read.csv("C:/Users/Hannah Bishop/Desktop/MeaS/Genes for in situ.csv",stringsAsFactors=FALSE)

# all.genes <- rownames(all) # same as rownames(all.tree)
all.genes <- rownames(all@assays$RNA) 

genelist <- vector()
nomatch <- list()
for (i in 1:12){
  gene <- yfg[[i,2]]
  #gene<-substr(gene,7,str_length(gene)-1)
  loc <- grep(gene,all.genes)
  if (length(loc)>0) {
    genelist <- c(genelist,loc)
  } else {
    nomatch <- c(nomatch,yfg[[i,2]])
  }
}

DefaultAssay(all.tree) <- "RNA"

DotPlot(all.tree,features=rev(all.genes[genelist[1:12]])) + RotatedAxis()

FeaturePlot(all.tree,features = all.genes[genelist[1:2]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree,features = all.genes[genelist[3:4]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree,features = all.genes[genelist[5:6]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree,features = all.genes[genelist[7:8]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree,features = all.genes[genelist[9:10]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree,features = all.genes[genelist[11:12]], ncol = 1) + NoLegend() + NoAxes()

VlnPlot(all.tree,features = all.genes[genelist[1:4]], ncol = 1) + NoLegend() + NoAxes()
VlnPlot(all.tree,features = all.genes[genelist[5:8]], ncol = 1) + NoLegend() + NoAxes()
VlnPlot(all.tree,features = all.genes[genelist[9:12]], ncol = 1) + NoLegend() + NoAxes()

# Can also use line of code below to find candidate gene 
#FeaturePlot(all.tree,features = all.genes[grep("VGlut",all.genes)])