---
title: "CCA integration of v3 datasets with HiSat GTF"
author: "jos"
date: "05/28/2020"
output:
  html_document:
    fig_height: 12
    fig_width: 17
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Objective: use a for loop to read in and process all 3 data sets from the original GTF and conduct a straight merge with standard normalization.
```{r, echo = FALSE}
# to knit document, use the following command: rmarkdown::render("C:/Users/Hannah Bishop/Desktop/GitHubCode/Octopolis/OSall_30pcs30dimsres1_normCCA_HiSat.Rmd") ~30m

library(Seurat)
library(Matrix)
library(ggplot2)
library(sctransform)
library(stringr)
library(cowplot) # used for CCA
library(patchwork) # used for CCA
```

## Read in data
```{r , echo = FALSE, eval = FALSE}
all_list <- readRDS("C:/Users/Hannah Bishop/Desktop/MeaS/hisatAll_list.rds")
for (i in 1:length(all_list)){
  #all_list[[i]][["percent.mt"]] <- PercentageFeatureSet(all_list[[i]], pattern = "^mitochondria-") #preferred for for loop over original command, leave out for new GTF because poor annotation
  all_list[[i]] <- NormalizeData(all_list[[i]])
  all_list[[i]] <- FindVariableFeatures(all_list[[i]], selection.method = "vst", nfeatures = 2000)
}

allnorm.anchors <- FindIntegrationAnchors(all_list[2:3], dims = 1:30) # runs CCA ~30m; shows 0% but moves on to merging objects etc; found 43006 anchors -> 10257 [what is the ideal/typical number for anchors across three datasets?]; overall takes ~2-3hr
allnorm.integrated <- IntegrateData(anchorset = allnorm.anchors, dims = 1:30)

saveRDS(allnorm.integrated, "C:/Users/Hannah Bishop/Desktop/MeaS/Seurat/Hisat/v3normCCA_Hisat.rds")
```

## Read in integrated dataset
I performed normalization on the individual datasets before conducting a merge using CCA. I applied normalization, scaledata, and findvariablefeatures on individual datasets before integrating based on anchors. I then saved this Seurat object so that I can easily read in the rds file when knitting the html.
```{r, echo = FALSE}
all <- readRDS("C:/Users/Hannah Bishop/Desktop/MeaS/Seurat/Hisat/v3normCCA_Hisat.rds")
DefaultAssay(all) <- "integrated"
```

## Plot features of the object. 
Visualization of object features: plot nFeature, nCount, and percent.mt for each of the Runs in the merged object. (Note: nFeature = # genes, nCount = UMI counts, and percent.mt is important because higher levels of percent.mt are indicative of cell stress.) Typically, I would subset the data to only include nFeatures >200, <2500, and percent.mt <5. However, even though I used visualization techniques to plot these features, I did not end up subsetting the data.
```{r }
VlnPlot(all, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
#plot1 <- FeatureScatter(all, feature1 = "nCount_RNA", feature2 = "percent.mt")
# visualize the relationship between UMI counts and mitochondrial genes (percent)
plot2 <- FeatureScatter(all, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") 
# visualize the relationship between genes (ID as >1 UMI) and mitochondrial genes
#plot1 + plot2
plot2
```

## Run PCA and generate UMAP
```{r }
all <- ScaleData(all)
all <- RunPCA(all, npcs = 30, verbose = FALSE) 
all <- FindNeighbors(all, reduction = "pca", dims = 1:30) #not sure if "reduction" argument is necessary
all <- FindClusters(all, resolution = 1)
all <- RunUMAP(all, reduction = "pca", dims = 1:30)
DimPlot(all, reduction = "umap", label = TRUE)
DimPlot(all, reduction = "umap", group.by = "orig.ident")
```

## Examine cell-cluster information
```{r }
table(Idents(all))
table(all$orig.ident)
prop.table(table(Idents(all)))
table(Idents(all), all$orig.ident)
```
## Build a hierarchical tree
```{r }
all.tree <- BuildClusterTree(all, reorder = TRUE, reorder.numeric = TRUE, slot = "scale.data", verbose = TRUE, dims = 1:30)
PlotClusterTree(all.tree, label = TRUE)
```

```{r }
all.tree <- RunUMAP(all.tree, dims = 1:30)
DimPlot(all.tree, reduction = "umap", label = TRUE)
table(Idents(all.tree)) 
table(all.tree$orig.ident) 
prop.table(table(Idents(all.tree))) 
table(Idents(all.tree), all.tree$orig.ident) 
```

## Plot features
````{r, echo = FALSE}
yfg <- read.csv("C:/Users/Hannah Bishop/Desktop/MeaS/Genes for in situ.csv",stringsAsFactors=FALSE)

# all.genes <- rownames(all) # same as rownames(all.tree)
all.genes <- rownames(all@assays$RNA) 

genelist <- vector()
nomatch <- list()
for (i in 1:12){
  gene <- yfg[[i,2]]
  #gene<-substr(gene,7,str_length(gene)-1)
  loc <- grep(gene,all.genes)
  if (length(loc)>0) {
    genelist <- c(genelist,loc)
  } else {
    nomatch <- c(nomatch,yfg[[i,2]])
  }
}

# hisat file was not able to find the following: 
# "Ocbimv22021175" = synaptotagmin OCTOGene.20163
# "Ocbimv22024718" = VGlut 
# "Ocbimv22017369" = TH OCTOGene.16681
# "Ocbimv22001324" = FMRF-amide OCTOGene.28623
# "Ocbimv22031264" = TyrBH OCTOGene.41051

DefaultAssay(all.tree) <- "RNA"

DotPlot(all.tree,features=rev(all.genes[genelist[1:7]])) + RotatedAxis()
DotPlot(all.tree, features = c("OCTOGene.20163", "OCTOGene.16681", "OCTOGene.28623", "OCTOGene.41051") #synaptotagmin, TH, FMRF-amide, TyrBH
)

FeaturePlot(all.tree,features = all.genes[genelist[1:2]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree,features = all.genes[genelist[3:4]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree,features = all.genes[genelist[5:7]], ncol = 1) + NoLegend() + NoAxes()
FeaturePlot(all.tree, features = c("OCTOGene.20163", "OCTOGene.16681", "OCTOGene.28623", "OCTOGene.41051")
)

VlnPlot(all.tree,features = all.genes[genelist[1:4]], ncol = 1) + NoLegend() + NoAxes()
VlnPlot(all.tree,features = all.genes[genelist[5:7]], ncol = 1) + NoLegend() + NoAxes()
VlnPlot(all.tree, features = c("OCTOGene.20163", "OCTOGene.16681", "OCTOGene.28623", "OCTOGene.41051")
)

# Can also use line of code below to find candidate gene 
#FeaturePlot(all.tree,features = all.genes[grep("VGlut",all.genes)])