---
title: "Reading in Data, Standard Norm of HiSeq with updated names 09/18/20 for Hannah & Elena"
author: "jos"
date: "9/17/2020"
output: 
  html_document:
    fig_height: 12
    fig_width: 17
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r, echo = FALSE}
# to knit document, use the following command: rmarkdown::render("/Users/josongco/GitHub/Octopolis/Std_CCA_HE.Rmd") ~30m [rename file when saved so code doesn't overwrite while producing next html]


library(Seurat)
library(Matrix)
library(ggplot2)
library(sctransform)
library(stringr)
library(cowplot) # used for CCA
library(patchwork) # used for CCA
library(dplyr) # used for print markers

```

## Set pathway & parameters
```{r}
# change depending on user
user = "mea"
if (user == "mea"){
  directoryMaster = "~/Google Drive/miller niell octo seq/project data/scRNAseq/Seurat/directoryMaster_070720.csv"
  full_namekey = "~/Google Drive/miller niell octo seq/project data/scRNAseq/Seurat/gene lists/geneIDs_miniNCBI_fulldf_namekey_Sep18_2.csv"
} else if (user == "gabby"){
  directoryMaster = "directoryMaster_070720.csv"
  full_namekey = "fulldf_namekey.csv"
}
```

## Read in files
```{r echo = FALSE, eval = FALSE}
# read in csv with the directories
datafiles <- read.csv(directoryMaster, stringsAsFactors = FALSE) 
data_dir <- as.vector(datafiles$Directory)

# read in csv namekey
namekey <- read.csv(full_namekey, stringsAsFactors = FALSE)
```

## Hashmap for replacing names: Original or Hisat GTF?
```{r echo = FALSE, eval = FALSE}
#making a hashmap of OctoGenes/Ocbims to fullnames from the seurat namekey 
gtf = "hisat"
if (gtf == "original") { # couldn't get this to work
    unclass(by(namekey, namekey$Ocbimv, function(x) {
     tmp <- x$fullname
      setNames(tmp, x$Ocbimv[1])
     tmp
     })) -> namekeyDict
} else if (gtf == "hisat"){
   unclass(by(namekey, namekey$OctoGene, function(x) {
     tmp <- x$fullname
     setNames(tmp, x$OctoGene[1])
      tmp
      })) -> namekeyDict
}
```

## Check that hashmap worked properly
```{r echo = FALSE, eval = FALSE}
if (gtf == "original") {
#Check that the hashmap is working properly (original gtf)
print(namekeyDict[["Ocbimv22027788m"]]) #should print  "NA-OG10040-OB27788"
print(namekeyDict[["Ocbimv22030350m"]]) #should print "Endonuclease-Exonuclease-phosphatase-family-phosphodiesterase-12-OG10002-OB30350" 
} else if (gtf == "hisat") {
#Check that the hashmap is working properly (hisat)
print(namekeyDict[["OCTOGene.1"]]) #should print "NA-OG1"
print(namekeyDict[["OCTOGene.10032"]]) #should print "anaphase-promoting-complex-subunit-7-OG10032-OB20676-OB29021"
}
```

## An overview of data processing is as follows:
1. Read in raw data
2. Replace names with gene IDS
3. Create a Seurat object for downstream analyses and data manipulation

# Using a for loop, load in Octo Seq Data and take data through pre-processing steps
```{r echo = FALSE, eval = FALSE}
#make empty list for all the seurat objects
all_list = list() 

ct = 1 #what is ct?
for(f in 1:2){ #Old or New gtf, NextSeq + HiSeq from directory master
  all <- Read10X(data.dir = data_dir[f+16]) # f+16 for new gtf HiSeq, f+10 for old gtf HiSeq
  ngenes <- length(all@Dimnames[[1]])
    if (gtf == "original"){ #replacing Ocbim names
      for (g in 1:ngenes){
        gene <- all@Dimnames[[1]][g]
        gene <- substr(gene, 6, str_length(gene)-2)
        ind<- grep(gene, namekey[[3]])
        if (length(ind)>0) {
          id <- namekey[[ind[1], 2]]
          if (str_length(id)>0){
            id <- str_remove_all(id, "\\(") # parentheses mess up gene names as dimensions
            id <- str_remove_all(id, "\\)") 
            id <- substr(id, 1, 60) # keep it short
            all@Dimnames[[1]][g] <- temp
        }
      }
        } 
      } else if (gtf == "hisat"){ #replacing Octogene names
      for (g in 1:ngenes){
        gene <- all@Dimnames[[1]][g]
        temp <- namekeyDict[[gene]]
        if (length(temp) == length(all@Dimnames[[1]][g])){ #This checks that temp is not zero
         all@Dimnames[[1]][g] <- temp
         }
       }
      }
  all <- CreateSeuratObject(counts = all, project = paste("Run", f, sep = ""), min.cells = 3, min.features = 750)
  all_list[[ct]] = all
  ct = ct + 1
}


# Check that all names were replaced correctly # does not work for original gtf
grep("VGlut",all_list[[1]]@assays[["RNA"]]@data@Dimnames[[1]], ignore.case = TRUE) # prints [1] 2646 5437
# double check that this is the right gene that you were looking for
all_list[[1]]@assays[["RNA"]]@data@Dimnames[[1]][5437] # prints "VGlut-Major-Facilitator-Superfamily-solute-carrier-family-17-sodium-dependent-inorganic-phosphate-cotransporter-member-6-OG8038-OB24718"

saveRDS(all_list, "/Users/josongco/Desktop/FinalSeqData/hisatAll_list_HE.rds") #change depending on dataset
  
# To access individual objects: all_list[[1]]
#example: all_list[[1]] will print the object information on Seurat object 1
```

```{r, echo = FALSE, eval = FALSE}
##### HISAT GTF
all_list <- readRDS("/Users/josongco/Desktop/FinalSeqData/hisatAll_list_HE.rds") #HiSeqCombined updated gene names 09/18/2020, for Hannah and Elena

for (i in 1:length(all_list)){
  #all_list[[i]][["percent.mt"]] <- PercentageFeatureSet(all_list[[i]], pattern = "^mitochondria-") #preferred for for loop over original command
   all_list[[i]] <- NormalizeData(all_list[[i]])
   all_list[[i]] <- FindVariableFeatures(all_list[[i]], selection.method = "vst") #omit nfeatures = 2000 for 9120, 9320, 91320, 91820, no names dataset
}

## Identify anchors and integrate datasets.
all.anchors <- FindIntegrationAnchors(object.list = all_list, anchor.features = 3000, verbose = TRUE) # can omit dims = 1:30 but takes a really long time to run; determine if want to keep anchor.features = 2000; anchor.features = 3000 for 91320 data; omit ndims for 9120, 9320, 91320, 91820 data 
allnorm.integrated <- IntegrateData(anchorset = all.anchors, verbose = TRUE)

saveRDS(allnorm.integrated, "/Users/josongco/Desktop/FinalSeqData/Hisat_HiSeqCombined_HE_StdCCA.rds") #omit nfeatures = 2000, ndims = 1:30 for nonames dataset
```

```{r, echo = FALSE}
PC = 30
DM = 1:30
DMset = 30
```

```{r echo = FALSE}
# functions to trim labels
# these could be merged to be one function with multiple options
# and maybe do something more interesting like trim post OG
# each takes in a ggplot variable and max label length to trim to
# returns the updated plot variable

# trim labels on x-axis
trimLabelsX <- function(p,len){
  pp<- ggplot_build(p)
  oldxlabels = pp$layout$panel_params[[1]]$x.labels
  if (length(oldxlabels)==0){
  oldxlabels = pp$layout$panel_params[[1]]$x$breaks
  }
  newxlabels = oldxlabels
  for (i in 1:length(oldxlabels)){
    i
    newxlabels[[i]] =  substr(oldxlabels[[i]],1,len)
  }
 p<- p + scale_x_discrete(labels = newxlabels)
}

#trim labels on y-axies
trimLabelsY <- function(p,len){
  pp<- ggplot_build(p)
  
  oldylabels = ggplot_build(p)$layout$panel_params[[1]]$y.labels
    if (length(oldylabels)==0){
  oldylabels = pp$layout$panel_params[[1]]$y$breaks
  }
  newylabels = oldylabels
for (i in 1:length(oldylabels)){
    newylabels[[i]] =  substr(oldylabels[[i]],1,len)
  }
  p + scale_y_discrete(labels = newylabels)
}

trimTitle <- function(p,len){
  title = ggplot_build(p)$plot$labels$title
  newtitle = substr(title,1,len)
  p + labs(title=newtitle)
}
```



```{r, echo = FALSE}
all <- readRDS("/Users/josongco/Desktop/FinalSeqData/Hisat_HiSeqCombined_HE_StdCCA.rds")

DefaultAssay(all) <- "integrated" #omit if using straight merge
```

```{r echo = FALSE}
VlnPlot(all, features = c("nFeature_RNA", "nCount_RNA"), pt.size = 0, ncol = 2)
VlnPlot(all, features = "nCount_RNA", y.max = 20000, pt.size = 0.0001, ncol = 3)
#plot1 <- FeatureScatter(all, feature1 = "nCount_RNA", feature2 = "percent.mt")
# visualize the relationship between UMI counts and mitochondrial genes (percent)
plot2 <- FeatureScatter(all, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
# visualize the relationship between genes (ID as >1 UMI) and mitochondrial genes
#plot1 + plot2
plot2
```

```{r, echo = FALSE, eval = FALSE } 
## Subset nFeature > 500
all <- subset(all, subset = nFeature_RNA > 500 & nFeature_RNA < 7000 & nCount_RNA <30000)
# eval features after subsetting: 
VlnPlot(all, features = c("nFeature_RNA", "nCount_RNA"), pt.size = 0, ncol = 2)
```

```{r echo = FALSE, eval = TRUE}
## Subset nFeature > 750
all <- subset(all, subset = nFeature_RNA > 750 & nFeature_RNA < 7000 & nCount_RNA <30000)
# eval features after subsetting: 
VlnPlot(all, features = c("nFeature_RNA"), pt.size = 0, y.max = 7500) + NoLegend()
VlnPlot(all, features = c("nCount_RNA"), pt.size = 0, y.max = 30000) + NoLegend() 
```

## Run PCA and generate UMAP
```{r echo = FALSE}
all <- ScaleData(all)
all <- RunPCA(all, npcs = PC, verbose = FALSE)
for (i in DM){ 
p<-DimHeatmap(all, dims = i, cells = 500, balanced = TRUE)
print(p)
}
#DimHeatmap(all, dims = 1:5, cells = 500, balanced = TRUE)
#DimHeatmap(all, dims = 6:10, cells = 500, balanced = TRUE)
#DimHeatmap(all, dims = 11:15, cells = 500, balanced = TRUE)
#DimHeatmap(all, dims = 16:20, cells = 500, balanced = TRUE)
#DimHeatmap(all, dims = 21:25, cells = 500, balanced = TRUE)
#DimHeatmap(all, dims = 26:30, cells = 500, balanced = TRUE)

ElbowPlot(all, ndims = DMset)

#all <- JackStraw(all, dims = DMset) # dims max depends on how many pcs originally calculated
#all <- ScoreJackStraw(all, dims = DM)
#JackStrawPlot(all, dims = DM)
```

```{r echo = FALSE}
# original clustering
all <- FindNeighbors(all, dims = DM) 
all <- FindClusters(all, resolution = 1)
all <- RunUMAP(all, reduction = "pca", dims = DM)
DimPlot(all, reduction = "umap", label = TRUE)
DimPlot(all, reduction = "umap", group.by = "orig.ident")
```

## Build a hierarchical tree & plot UMAP
```{r echo = FALSE}
all.tree <- BuildClusterTree(all, reorder = TRUE, reorder.numeric = TRUE, slot = "scale.data", verbose = TRUE, dims = DM)
PlotClusterTree(all.tree, label = TRUE)
DimPlot(all.tree, reduction = "umap", label = TRUE)
DimPlot(all.tree, reduction = "umap", group.by = "orig.ident")
saveRDS(all.tree, "/Users/josongco/Desktop/FinalSeqData/alltree_HE.rds")
```

## Examine cell-cluster information & cluster features
```{r echo = FALSE}
id_cell <- table(Idents(all.tree))
barplot(id_cell, main = "cell # in each cluster", xlab = "cluster ID")
#table(all.tree$orig.ident)
#prop.table(table(Idents(all.tree)))
id_runs <- table(all.tree$orig.ident, Idents(all.tree))
#barplot(id_runs, main = "relationship between cluster ID and run representation", xlab = "cluster ID", ylab = "cells", col = c("lightcoral", "palegreen3", "steelblue1"), legend = rownames(id_runs), beside=TRUE) # all runs
barplot(id_runs, main = "relationship between cluster ID and run representation", xlab = "cluster ID", ylab = "cells", col = c("lightcoral", "steelblue1"), legend = rownames(id_runs), beside=TRUE) # v3 chem

plot(id_runs[1,],id_runs[2,],xlab = "Run 1",ylab="Run 2")

VlnPlot(all.tree, features = "nCount_RNA", y.max = 20000, pt.size = 0, ncol = 1) + NoLegend()
VlnPlot(all.tree, features = "nFeature_RNA", y.max = 7500, pt.size = 0, ncol = 1) + NoLegend()
```

## ID cluster markers based on dendogram renumbering
```{r echo = FALSE} 
markers_list = list()
for (m in 1:nlevels((all.tree$seurat_clusters))){
  markersROC <- FindMarkers(all.tree, ident.1 = m, logfc.threshold = 0.5, test.use = "roc", only.pos = TRUE)
  markersROC <- setNames(cbind(rownames(markersROC), markersROC, row.names = NULL), c("geneID", "myAUC", "avg_diff", "power", "pct.1", "pct.2"))
  markersROC$diff_exp <- (markersROC$pct.1/markersROC$pct.2)
  markersROC <- markersROC[order(-markersROC$power),]
  header <- paste("Cluster", m, sep = " ")
  #top5markers <- paste((head(markersROC$geneID, n = 5)))
  #marker_output <- c(header, top5markers)
  #print(markersROC)
  markers_list[m] = markersROC
  
  p<-DotPlot(all.tree,features=rev(head(markersROC$geneID,50))) + RotatedAxis()
p<- trimLabelsX(p,20)
p<- p + labs(title = header)
print(p)

}

# write.csv(markers_list, "/Users/josongco/Google Drive/miller niell octo seq/project data/scRNAseq/Seurat/clustermarkers9120.csv") #9120 indicates dataset
```

## ID markers for nodes/connecting points of dendogram
```{r echo = FALSE }
# works but sometimes crashes - too big?
print(PlotClusterTree(all.tree, label = TRUE))

nodes <- unique(all.tree@tools$BuildClusterTree$edge[,1])
node_markers = list()
for (n in 1:length(nodes)){
    nodemarkersROC <- FindMarkers(all.tree, ident.1 = "clustertree", ident.2 = nodes[n],  test.use = "roc", only.pos = TRUE)
  nodemarkersROC <- setNames(cbind(rownames(nodemarkersROC), nodemarkersROC, row.names = NULL), c("geneID", "myAUC", "avg_diff", "power", "pct.1", "pct.2"))
  nodemarkersROC$diff_exp <- (nodemarkersROC$pct.1/nodemarkersROC$pct.2)
  leftMarkers <- nodemarkersROC[nodemarkersROC$avg_diff>0,]
  leftMarkers <- leftMarkers[order(-leftMarkers$power),]
  rightMarkers <- nodemarkersROC[nodemarkersROC$avg_diff<0,]
  rightMarkers <- rightMarkers[order(-rightMarkers$power),]
  
  leftrightMarkers = rbind(head(leftMarkers,25),head(rightMarkers,25))
  nodemarkersROC <- nodemarkersROC[order(-nodemarkersROC$power),]
  node_markers[[n]] = leftrightMarkers
  node_header <- paste("Node", nodes[n], sep = " ")
  #top5nodemarkers <- paste(head(nodemarkersROC$geneID, n = 5))
  #nodemarker_output <- c(node_header, top5nodemarkers)
 #print(nodemarker_output)
  
#  p<-DotPlot(all.tree,features=rev(head(nodemarkersROC$geneID,50))) + RotatedAxis()
#p<- trimLabelsX(p,20)
#p<- p + labs(title = node_header)
#print(p)

  p<-DotPlot(all.tree,features=rev(leftrightMarkers$geneID)) + RotatedAxis()
p<- trimLabelsX(p,20)
p<-p + labs(title = node_header)
print(p)

}
```

## Plot features
````{r, echo = FALSE, eval = TRUE}
### function to find list of ocbims in all gene names
### takes a list of ocbims, returns matching index for each one
all.genes <- rownames(all.tree@assays$RNA) 

findGenes <- function(yfg,allgenes) {
  genelist <- vector()
  nomatch <- list()
  for (i in 1:length(yfg)){
    gene <- yfg[[i]]
    gene<-substr(gene,10,str_length(gene))
    gene<- paste("OB", gene, sep = "")
    loc <- grep(gene,allgenes)
    # check that there was one identified, and that the gene wasn't empty (will be OB and match everything!)
    if (length(loc)>0& nchar(gene)>2) {
      genelist <- c(genelist,loc)
    } else {
      nomatch <- c(nomatch,yfg[[i]])
    }
  }
  return(genelist)
}

```

```{r echo = FALSE, eval = TRUE}
yfg <- read.csv("~/Google Drive/miller niell octo seq/project data/scRNAseq/Seurat/gene lists/Genes for in situ.csv",stringsAsFactors=FALSE)

#yfg <- read.csv("D:/data/octo seq/Genes for in situ.csv",stringsAsFactors=FALSE)

# find match for yfg and all gene names

genelist <- findGenes(yfg[1:12,2],all.genes)


# hisat file was not able to find the following: 
# "Ocbimv22021175" = synaptotagmin OG.20163
# "Ocbimv22024718" = VGlut OG.8038
# "Ocbimv22017369" = TH OG.16681
# "Ocbimv22001324" = FMRF-amide OG.28623
# "Ocbimv22031264" = TyrBH OG.41051

DefaultAssay(all.tree) <- "RNA"

# dotplot of all genes from list
p<-DotPlot(all.tree,features=rev(all.genes[genelist])) + RotatedAxis()
p<- trimLabelsX(p,20)
print(p)

# featureplot of all genes in list
for (i in 1:length(genelist)){
  p<- FeaturePlot(all.tree, features = all.genes[genelist[i]], ncol = 1) + NoLegend() + NoAxes() 
  p<- trimTitle(p,30)
  print(p)
}

# violin plot for all genes in list
for (i in 1:length(genelist)){
  p <- VlnPlot(all.tree, features = all.genes[genelist[i]], ncol = 1, pt.size = 0) + NoLegend() + NoAxes()
  p<-trimTitle(p,30)
  print(p)
}

#dotplot of some selected genes
#DotPlot(all.tree, features = c("OG.20163", "OG.16681", "OG.28623", "OG.41051", "OG.8038")) #synaptotagmin, TH, FMRF-amide, TyrBH, VGlut)

#FeaturePlot(all.tree, features = c("OG.20163", "OG.16681", "OG.28623", "OG.41051", "OG.8038"), ncol = 1)
#FeaturePlot(all.tree, features = ("VACHT"))


#VlnPlot(all.tree, features = c("OG.20163", "OG.16681", "OG.28623", "OG.41051", "OG.8038"), pt.size = 0, ncol = 1)

#grep("glut",all.genes)
 #[1]  2034  4008  4599 10227 12291 14312 18187 19855 19948 27719 32028 40993 41561

#VlnPlot(all.tree,features = c(all.genes[2034],all.genes[4008], all.genes[4599], all.genes[10227]), pt.size = 0, ncol = 1) 
#VlnPlot(all.tree, features = c(all.genes[12291], all.genes[14312], all.genes[18187], all.genes[19855]), pt.size = 0, ncol = 1) 
#VlnPlot(all.tree, features  = c(all.genes[19948], all.genes[27719], all.genes[32028], all.genes[40993], all.genes[41561]), pt.size = 0, ncol = 1) # glut

# Can also use line of code below to find candidate gene 
#FeaturePlot(all.tree,features = all.genes[grep("VGlut",all.genes)])
#FeaturePlot(all.tree,features = all.genes[grep("glut",all.genes)]) * use this one!!
#
#1synaptotagmin	Ocbimv22021175	NA	neurons	NA	G
#2	VAchT	Ocbimv22001681	NA	cholinergic neurons	NA	R
#3	ChaT	Ocbimv22001674	NA	cholinergic neurons	NA	G
#4	VGlut	Ocbimv22024718	NA	glutamatergic neurons	NA	R ------
#5	TH	Ocbimv22017369	NA	dopaminergic neurons (and pigment?)	NA	R
#6	DAT	Ocbimv22026818	NA	just dopaminergic?	NA	G
#7	VMAT-A	Ocbimv22031489	NA	monoamine transporter	NA	G ------
#8	FMRF-related peptide	Ocbimv22023842	NA	peptide transmission	NA	R
#9	FMRF-amide	Ocbimv22001324	NA	distinct from FRMF-related peptide	NA	R
#10	TyrBH	Ocbimv22031264	NA	Octopamine	NA	R
#11	AchE	Ocbimv22038398	NA	glia?	NA	R
#12	GAD
#FeaturePlot(all.tree,features = all.genes[grep("03827",all.genes)]) 
```


```{r echo = FALSE, eval = FALSE}
# grep for genes
genelist = grep("VGlut",all.genes, ignore.case = TRUE)

# dotplot
p<-DotPlot(all.tree,features=rev(all.genes[genelist])) + RotatedAxis()
p<- trimLabelsX(p,20)
print(p)

#feature plots

for (i in 1:length(genelist)){
  p<- FeaturePlot(all.tree, features = all.genes[genelist[i]], ncol = 1) + NoLegend() + NoAxes() 
  p<- trimTitle(p,30)
  print(p)
}

```

