---
title: "TopoMapp_ObimacVS_Analysis"
author: "JO Songco-Casey"
date: "2024-08-19"
output: html_document
---

Mount to talapas to access raw cellranger files:
sshfs jsongco@login.talapas.uoregon.edu:/projects/niell/shared/Octopus /Users/josongco/Octopus

```{r}
# set working directory in console
setwd("/Users/josongco/Octopus") # when reading in talapas files
setwd('/Users/josongco/Desktop/topographicalMapping') # when saving files locally
```

Begin 'Integrative analysis in Seurat v5' Vignette
https://satijalab.org/seurat/articles/seurat5_integration

```{r}
library(Seurat)
library(SeuratData)
library(SeuratWrappers)
library(Azimuth)
library(ggplot2)
library(patchwork)
```

```{r}
# read in files
directoryMaster = "/Users/josongco/Desktop/topographicalMapping/directoryMaster.csv"
datafiles <- read.csv(directoryMaster, stringsAsFactors = FALSE)
data_dir <- as.vector(datafiles$Directory)

#make empty list for all the seurat objects
all_list = list()

ct = 1

for(f in 1:4){
  all <- Read10X(data.dir = (data_dir[f]))
  all <- CreateSeuratObject(counts = all, project = paste("Sample", f, sep = ""), min.cells = 3, min.features = 500)
all_list[[ct]] = all
  ct = ct + 1
}

saveRDS(all_list, "/Users/josongco/Desktop/topographicalMapping/datasets/completeLRTB.rds") #change depending on dataset
```

```{r}
# identify mitochondrial genes - included in genome without "obimac" identifier

for (i in 1:length(all_list)){
  all_list[[i]][["percent.mt"]] <- PercentageFeatureSet(all_list[[i]], pattern = "^[^obimac]")
}

# visualize reads per cell
Thres_nCount_1 <- VlnPlot(all_list[[1]], features = "nCount_RNA", y.max = 20000, pt.size = 0, cols = "lightcoral")
Thres_nCount_2 <- VlnPlot(all_list[[2]], features = "nCount_RNA", y.max = 20000,pt.size = 0, cols = "steelblue") 
Thres_nCount_3 <- VlnPlot(all_list[[3]], features = "nCount_RNA", y.max = 20000,pt.size = 0, cols = "lightgreen")
Thres_nCount_4 <- VlnPlot(all_list[[4]], features = "nCount_RNA", y.max = 20000,pt.size = 0, cols = "yellow") 
wrap_plots(Thres_nCount_1, Thres_nCount_2, Thres_nCount_3, Thres_nCount_4)
Thres_nCount_1 + Thres_nCount_2 + Thres_nCount_3 + Thres_nCount_4

Thres_nCount_1 <- VlnPlot(all_list[[1]], features = "nCount_RNA", y.max = 10000, pt.size = 0, cols = "lightcoral")
Thres_nCount_2 <- VlnPlot(all_list[[2]], features = "nCount_RNA", y.max = 10000,pt.size = 0, cols = "steelblue") 
Thres_nCount_3 <- VlnPlot(all_list[[3]], features = "nCount_RNA", y.max = 10000,pt.size = 0, cols = "lightgreen")
Thres_nCount_4 <- VlnPlot(all_list[[4]], features = "nCount_RNA", y.max = 10000,pt.size = 0, cols = "yellow") 
wrap_plots(Thres_nCount_1, Thres_nCount_2, Thres_nCount_3, Thres_nCount_4)

Thres_nCount_1 <- VlnPlot(all_list[[1]], features = "nCount_RNA", y.max = 2500, pt.size = 0, cols = "lightcoral")
Thres_nCount_2 <- VlnPlot(all_list[[2]], features = "nCount_RNA", y.max = 2500,pt.size = 0, cols = "steelblue") 
Thres_nCount_3 <- VlnPlot(all_list[[3]], features = "nCount_RNA", y.max = 2500,pt.size = 0, cols = "lightgreen")
Thres_nCount_4 <- VlnPlot(all_list[[4]], features = "nCount_RNA", y.max = 2500,pt.size = 0, cols = "yellow") 
wrap_plots(Thres_nCount_1, Thres_nCount_2, Thres_nCount_3, Thres_nCount_4)

# visualize genes per cell
Thres_nFeat_1 <- VlnPlot(all_list[[1]], features = "nFeature_RNA", y.max = 5000,pt.size = 0, cols = "lightcoral") 
Thres_nFeat_2 <- VlnPlot(all_list[[2]], features = "nFeature_RNA", y.max = 5000,pt.size = 0, cols = "steelblue") 
Thres_nFeat_3 <- VlnPlot(all_list[[3]], features = "nFeature_RNA", y.max = 5000,pt.size = 0, cols = "lightgreen") 
Thres_nFeat_4 <- VlnPlot(all_list[[4]], features = "nFeature_RNA", y.max = 5000,pt.size = 0, cols = "yellow") 
wrap_plots(Thres_nFeat_1, Thres_nFeat_2, Thres_nFeat_3, Thres_nFeat_4)
Thres_nFeat_1+ Thres_nFeat_2+ Thres_nFeat_3+ Thres_nFeat_4

Thres_nFeat_1 <- VlnPlot(all_list[[1]], features = "nFeature_RNA", y.max = 1000,pt.size = 0, cols = "lightcoral") 
Thres_nFeat_2 <- VlnPlot(all_list[[2]], features = "nFeature_RNA", y.max = 1000,pt.size = 0, cols = "steelblue") 
Thres_nFeat_3 <- VlnPlot(all_list[[3]], features = "nFeature_RNA", y.max = 1000,pt.size = 0, cols = "lightgreen") 
Thres_nFeat_4 <- VlnPlot(all_list[[4]], features = "nFeature_RNA", y.max = 1000,pt.size = 0, cols = "yellow") 
wrap_plots(Thres_nFeat_1, Thres_nFeat_2, Thres_nFeat_3, Thres_nFeat_4)
Thres_nFeat_1+ Thres_nFeat_2+ Thres_nFeat_3+ Thres_nFeat_4

# visualize mitochondrial percentage per cell
Thres_MT_1 <- VlnPlot(all_list[[1]], features = "percent.mt", y.max = 20,pt.size = 0, cols = "lightcoral") 
Thres_MT_2 <- VlnPlot(all_list[[2]], features = "percent.mt",y.max = 20,pt.size = 0, cols = "steelblue") 
Thres_MT_3 <- VlnPlot(all_list[[3]], features = "percent.mt", y.max = 20,pt.size = 0, cols = "lightgreen") 
Thres_MT_4 <- VlnPlot(all_list[[4]], features = "percent.mt", y.max = 20,pt.size = 0, cols = "yellow") 
wrap_plots(Thres_MT_1 , Thres_MT_2, Thres_MT_3, Thres_MT_4)
Thres_MT_1+ Thres_MT_2+ Thres_MT_3+ Thres_MT_4

# set thresholds
for (i in 1:length(all_list)){
all_list[[i]] <- subset(all_list[[i]], subset = percent.mt < 6 & nCount_RNA > 1000 & nCount_RNA < 10000 & nFeature_RNA > 500)
}
```


Integrate datasets
```{r}
all.list <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/completeLRTB.rds")
#all.list = all_list
all.list <- lapply(X = all.list[3:4], FUN = function(x) {
    x <- NormalizeData(x)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})
features <- SelectIntegrationFeatures(object.list = all.list)
all.anchors <- FindIntegrationAnchors(object.list = all.list, anchor.features = features)
all.combined <- IntegrateData(anchorset = all.anchors)

saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/combinedLRTB.rds")
saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/combinedLR.rds")
saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/combinedTB.rds")

all.list <- lapply(X = all.list, FUN = SCTransform) 
features <- SelectIntegrationFeatures(object.list = all.list, nfeatures = 3000)

all.list <- PrepSCTIntegration(
  all.list,
  assay = "RNA",
  anchor.features = 3000,
  sct.clip.range = NULL,
  verbose = TRUE
)

all.list <- PrepSCTIntegration(object.list = all.list, anchor.features = features)

all.anchors <- FindIntegrationAnchors(object.list = all.list, anchor.features = features, normalization.method = "SCT", verbose = TRUE) 
allnorm.integrated <- IntegrateData(anchorset = all.anchors, normalization.method = "SCT", verbose = TRUE)

all.list <- SCTransform(all.list, vars.to.regress = "percent.mt", verbose = FALSE)


# these steps were performed using an earlier version of R and may require additional steps with later versions
 all_list <- lapply(X = all_list, FUN = SCTransform) 
 features <- SelectIntegrationFeatures(object.list = all_list, nfeatures = 3000)
 all_list <- PrepSCTIntegration(object.list = all_list, anchor.features = features)

all.anchors <- FindIntegrationAnchors(object.list = all_list, anchor.features = features, normalization.method = "SCT", verbose = TRUE) 
allnorm.integrated <- IntegrateData(anchorset = all.anchors, normalization.method = "SCT", verbose = TRUE)

# we recommend saving out the seurat object at this step in case R crashes  
saveRDS(allnorm.integrated, "/userpath/ObimacSeq_SCTCCAint.rds")
```

```{r}
saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/all.combined-std.rds")

all.combined <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/all.combined-std.rds")
all.combined <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/combinedLR.rds")
all.combined.sct <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/completeLR-IntegratedSct.rds")

DefaultAssay(all.combined.sct) <- "integrated"

# Run the standard workflow for visualization and clustering
all.combined <- ScaleData(all.combined, verbose = FALSE) # omit from sct datasets
all.combined <- RunPCA(all.combined.sct, npcs = 30, verbose = FALSE)

ElbowPlot(all.combined)

all.combined <- RunUMAP(all.combined, reduction = "pca", dims = 1:15)
all.combined <- FindNeighbors(all.combined, reduction = "pca", dims = 1:15)
all.combined <- FindClusters(all.combined, resolution = 0.5)


saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/umapLRTB.rds")
saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/umapTB.sct15d.rds")
saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/umapLR.sct15d.rds")
saveRDS(all.combined, "/Users/josongco/Desktop/topographicalMapping/datasets/umapLR.rds")

p1 <- DimPlot(all.combined, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(all.combined, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2

DimPlot(all.combined, reduction = "umap", split.by = "orig.ident", label = TRUE)
```

```{r}
#plotting genes on featureplots

all.combinedLR <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/umapLR.sct15d.rds")
all.combinedTB <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/umapTB.sct15d.rds")


pdf(file = "/Users/josongco/Desktop/topographicalMapping/latro-ten.pdf",
    width = 8, # The width of the plot in inches
    height = 8)
p1 <- DimPlot(all.combinedLR, label = TRUE) + ggtitle('left v right')
p2 <- DimPlot(all.combinedLR, group.by = "orig.ident")

DefaultAssay(all.combinedLR) <- "SCT"

fp1 <- FeaturePlot(all.combinedLR, features='obimac0021926')
fp2 <- FeaturePlot(all.combinedLR, features='obimac0020133')


p3 <- DimPlot(all.combinedTB, label = TRUE) + ggtitle('top v bottom')
p4 <- DimPlot(all.combinedTB, group.by = "orig.ident")

DefaultAssay(all.combinedTB) <- "SCT"

fp3 <- FeaturePlot(all.combinedTB, features='obimac0021926')
fp4 <- FeaturePlot(all.combinedTB, features='obimac0020133')



print(p1)
print(p2) 
print(fp1) 
print(fp2)


print(p3) 
print(p4) 
print(fp3)
print(fp4)


dev.off()


pdf(file = "/Users/josongco/Desktop/topographicalMapping/latro-ten.dp.pdf",
    width = 8, # The width of the plot in inches
    height = 4)

DefaultAssay(all.combinedLR) <- 'integrated'

DotPlot(all.combinedLR, features = c('obimac0021926','obimac0020133'), cols = c("blue", "red"), split.by = "orig.ident") + RotatedAxis() + coord_flip() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7)) + ggtitle('left v right, integrated assay')

DefaultAssay(all.combinedLR) <- 'SCT'

DotPlot(all.combinedLR, features = c('obimac0021926','obimac0020133'), cols = c("blue", "red"), split.by = "orig.ident") + RotatedAxis() + coord_flip() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7)) + ggtitle('left v right, sct assay')

DefaultAssay(all.combinedTB) <- 'integrated'

DotPlot(all.combinedTB, features = c('obimac0021926','obimac0020133'), cols = c("blue", "red"), split.by = "orig.ident") + RotatedAxis() + coord_flip() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7)) + ggtitle('top v bottom, integrated assay')

DefaultAssay(all.combinedTB) <- 'SCT'

DotPlot(all.combinedTB, features = c('obimac0021926','obimac0020133'), cols = c("blue", "red"), split.by = "orig.ident") + RotatedAxis() + coord_flip() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7)) + ggtitle('top v bottom, sct assay')



dev.off()

all.combined <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/umapLRTB.rds")
DefaultAssay(all.combined) <- "RNA"
FeaturePlot(all.combined, features = c('obimac0021926','obimac0020133'), split.by = "orig.ident")

```


Identify conserved cell type markers
```{r}
DefaultAssay(all.combined) <- "RNA"
all.combined <- JoinLayers(all.combined)
conserved.markers.1 <- FindConservedMarkers(all.combined, ident.1 = 1, grouping.var = "orig.ident", verbose = FALSE)
head(conserved.markers)

FeaturePlot(all.combined, features = c("G15407", "G13308", "G12130", "G14919", "G21142", "G14903"), min.cutoff = "q9", split.by = "orig.ident")
#immune.combined <- RenameIdents(immune.combined, `0` = "CD14 Mono", `1` = "CD4 Naive T", `2` = "CD4 Memory T", `3` = "CD16 Mono", `4` = "B", `5` = "CD8 T", `6` = "NK", `7` = "T activated", `8` = "DC", `9` = "B Activated", `10` = "Mk", `11` = "pDC", `12` = "Eryth", `13` = "Mono/Mk Doublets", `14` = "HSPC")
#DimPlot(immune.combined, label = TRUE)

#Idents(all.combined) <- "seurat_clusters"
DotPlot(all.combined, features = c("G15407", "G13308", "G12130", "G14919", "G21142", "G14903"), cols = c("blue", "red"), dot.scale = 8, split.by = "orig.ident") + RotatedAxis()
```

## set up functions for labels
```{r echo = FALSE, eval = TRUE}
# function to add gene names to data table
# assumes first column of table is G gene
# uses master list with gene name in 1st column, G gene in 3rd column
addGeneNamesTable<- function(data,geneID, gene, desc){ # a little redundant to send in geneID but this solves the problem that different objects name it differently
  data$GeneName <- '' # create empty column
  for (i in 1:nrow(data)){
    #ggene = data$gene[[i]] #cluster markers
    obgene = geneID[[i]] # individual cluster markers
    #ggene = data$Feature[[i]] #dimheatmap
    #ggene = data$feature[[i]] #viz dim loadings
    loc <- which(gene ==obgene)
    if (length(loc)>0) {  # if something is found
      data$GeneName[[i]] <- desc[loc[1]] # if multiple found, use 1st one
    }
  }

  data
}
#
#replaceLabelsX <- function(p,gene,desc){
#  pp<- ggplot_build(p)
#  oldxlabels = ggplot_build(p)$layout$panel_params[[1]]$x$breaks
#  newxlabels = oldxlabels
#  for (i in 1:length(oldxlabels)){
#    loc<-which(gene==(oldxlabels[[i]]))
#    if (length(loc)>0){
#      newxlabels[[i]] = paste(oldxlabels[[i]], "-", desc[[loc[1]]])
#      newxlabels[[i]] = substr(newxlabels[[i]],1,50)}
#    else {
#      for (j in 1:length(newxlabels)){
#        inds<-which(noID$ggene==(newxlabels[[j]]))
#        if (length(inds)>0){
#          newxlabels[[j]] = paste(newxlabels[[j]], "-", #noID$Manual_NCBI_BLAST_ID[[inds[1]]])
#        }
#        newxlabels[[j]] = substr(newxlabels[[j]],1,50)
#}}}
#  p<- p + scale_x_discrete(labels = newxlabels)
#}

# does not use noID list
replaceLabelsX <- function(p,gene,desc){
  pp<- ggplot_build(p)
  oldxlabels = ggplot_build(p)$layout$panel_params[[1]]$x$breaks
  newxlabels = oldxlabels
  for (i in 1:length(oldxlabels)){
    loc<-which(gene==(oldxlabels[[i]]))
    if (length(loc)>0){
      newxlabels[[i]] = paste(oldxlabels[[i]], "-", desc[[loc[1]]])
      newxlabels[[i]] = substr(newxlabels[[i]],1,50)}}
  p<- p + scale_x_discrete(labels = newxlabels)
}

replaceLabelsY <- function(p,gene,desc){
  pp<- ggplot_build(p)
  oldylabels = ggplot_build(p)$layout$panel_params[[1]]$y$breaks
  newylabels = oldylabels
  for (i in 1:length(oldylabels)){
    loc<-which(gene==(oldylabels[[i]])) # i/ind 4 = 919, G34789, NPcorrendocrine protein/ label = paste(genelabels[919], "-", namelist[919])
    if (length(loc)> 0) {
      #loc<-which(oldylabels %in% genelabels)
      newylabels[[i]] = paste(oldylabels[[i]], "-", desc[[loc[1]]])
      newylabels[[i]] = substr(newylabels[[i]],1,50)}
    else {
      for (j in 1:length(newylabels)){
        inds<-which(noID$ggene==(newylabels[[j]]))
        if (length(inds)>0){
          newylabels[[j]] = paste(newylabels[[j]], "-", noID$Manual_NCBI_BLAST_ID[[inds[1]]])
        }
        newylabels[[j]] = substr(newylabels[[j]],1,50)
      }}}
  p<- p + scale_y_discrete(labels = newylabels)
}


wrapper <- function(x, ...) 
{
  paste(strwrap(x, ...), collapse = "\n")
}

shorten <- function(x)
{
  stringr::str_trunc(x, 100)
}

gene_master <- read.csv("/Users/josongco/Desktop/Cell Types Seurat Code & Files/GeneIDs.csv")
```

Identifying differentially expressed genes across conditions
```{r}
library(ggplot2)
library(cowplot)
theme_set(theme_cowplot())

all.combined <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/umapTB.sct15d.rds")
all.combined <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/umapLR.rds")
all.combined <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/umapTB.rds")

cells1 <- subset(all.combined, idents = 1)
Idents(cells1) <- "orig.ident"
avg.cells1 <- as.data.frame(log1p(AverageExpression(cells1, verbose = FALSE)$SCT)) #$RNA
avg.cells1$gene <- rownames(avg.cells1)
avg.cells1 <- addGeneNamesTable(avg.cells1, avg.cells1$gene, avg.cells1$gene, gene_master$merge_desc)

cells2 <- subset(all.combined, idents = 2)
Idents(cells2) <- "orig.ident"
avg.cells2 <- as.data.frame(log1p(AverageExpression(cells2, verbose = FALSE)$SCT))
avg.cells2$gene <- rownames(avg.cells2)
avg.cells2 <- addGeneNamesTable(avg.cells2, avg.cells2$gene, avg.cells2$gene, gene_master$merge_desc)



genes.to.label = c("obimac0020958", "obimac0014441", "obimac0021555", "obimac0012076", "obimac0030402", "obimac0013580", "obimac0002637", "obimac0021237", "obimac0007601", "obimac0030399", "obimac0020553", "obimac0019980")
p1 <- ggplot(avg.cells1, aes(Sample1, Sample2)) + geom_point() + ggtitle("example cells idents 1")
p1 <- LabelPoints(plot = p1, points = genes.to.label, repel = TRUE)
p2 <- ggplot(avg.cells2, aes(Sample1, Sample2)) + geom_point() + ggtitle("example cells idents 2")
p2 <- LabelPoints(plot = p2, points = genes.to.label, repel = TRUE)
p1 + p2

allcombined.celltypesample <- all.combined
allcombined.celltypesample$celltype.sample <- paste(Idents(allcombined.celltypesample), allcombined.celltypesample$orig.ident, sep = "_")
allcombined.celltypesample$celltype <- Idents(allcombined.celltypesample)
Idents(allcombined.celltypesample) <- "celltype.sample"

DefaultAssay(all.combined) <- "integrated"
#example.combined <- JoinLayers(all.combined)
all.combined <- PrepSCTFindMarkers(all.combined)
exTB.markers <- FindConservedMarkers(all.combined, ident.1 = 5, assay = "SCT", grouping.var = "orig.ident", verbose = FALSE)
exTB.markers$gene <- rownames(exTB.markers)
exTB.markers <- addGeneNamesTable(exTB.markers, exTB.markers$gene, exTB.markers$gene, gene_master$merge_desc)


example.response <- FindMarkers(allcombined.celltypesample, ident.1 = "0_Sample3", ident.2 = "0_Sample4", verbose = FALSE)
head(example.response, n = 15)
example.response <- rownames(example.response)
example.response <- data.frame(example.response)
example.response$gene <- example.response$example.response
example.response <- addGeneNamesTable(example.response, example.response$gene, example.response$gene, gene_master$merge_desc)
top30_btwn <- head(example.response, n = 30)






genes.to.label <- read.csv("/Users/josongco/Desktop/topographicalMapping/mainGenes.csv")
genes.to.label <- addGeneNamesTable(genes.to.label, genes.to.label$obgene, gene_master$obgene, gene_master$merge_desc)
#DefaultAssay(all.combined) <- "RNA" # use for standard normalization
 
DefaultAssay(all.combined) <- "SCT" # use for sct normalization

pdf(file = "/Users/josongco/Desktop/topographicalMapping/LRmainGenesplots-sct15d.pdf",
    width = 8, # The width of the plot in inches
    height = 8)
  p1 <- DimPlot(all.combined, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(all.combined, reduction = "umap", label = TRUE, repel = TRUE)
print(p2)
print(p1)
for (features in 1:11){ #without no gene desc 
  fp <- FeaturePlot(all.combined, features = genes.to.label$obgene[features])
  fp <- fp + ggtitle(genes.to.label$obgene[features], subtitle = paste(genes.to.label$gene[features]))
  print(fp)
}

dots <- DotPlot(all.combined, features = genes.to.label$obgene)  + RotatedAxis() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7))
dots<-replaceLabelsX(dots,genes.to.label$obgene,genes.to.label$gene)
dots <- dots 
print(dots)
dev.off()

pdf(file = "/Users/josongco/Desktop/topographicalMapping/TBsct.conserved5.pdf",
    width = 8, # The width of the plot in inches
    height = 8)
DefaultAssay(all.combined) <- "SCT"
  p1 <- DimPlot(all.combined, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(all.combined, reduction = "umap", label = TRUE, repel = TRUE)
print(p2)
print(p1)
for (features in 1:20){ #without no gene desc 
  fp <- FeaturePlot(all.combined, features = exTB.markers$gene[features])
  fp <- fp + ggtitle(exTB.markers$gene[features], subtitle = paste(exTB.markers$GeneName[features]))
  print(fp)
}

dots <- DotPlot(all.combined, features = exTB.markers$gene[1:20], cols = c("blue","red"), split.by = "orig.ident")  + RotatedAxis() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7))
dots<-replaceLabelsX(dots,exTB.markers$gene,exTB.markers$GeneName)
dots <- dots 
print(dots)
dev.off()


FeaturePlot(all.combined, features = genes.to.label, split.by = "orig.ident", max.cutoff = 3, cols = c("grey", "red"))
VlnPlot(all.combined, features = c("G15407", "G13308", "G12130", "G14919"), split.by = "orig.ident", cols = c("grey", "red"), ncol = 4)
VlnPlot(all.combined, features = c("G15407", "G13308"), split.by = "orig.ident", cols = c("grey", "red"))

plots <- VlnPlot(all.combined, features = c("G15407", "G13308", "G12130"), split.by = "orig.ident", group.by = "celltype",
    pt.size = 0, combine = FALSE)
wrap_plots(plots = plots, ncol = 1)
```

Integration on datasets normalized with SCTransform
```{r}
all.list <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/completeLRTB.rds")
options(future.globals.maxSize = 8000 * 1024^2)
all.list <- lapply(X = all.list, FUN = SCTransform)
features <- SelectIntegrationFeatures(object.list = all.list, nfeatures = 3000)
all.list <- PrepSCTIntegration(object.list = all.list, anchor.features = features)
all.anchors <- FindIntegrationAnchors(object.list = all.list, normalization.method = "SCT",
    anchor.features = features)
all.combined.sct <- IntegrateData(anchorset = all.anchors, normalization.method = "SCT")

saveRDS(all.combined.sct, "/Users/josongco/Desktop/topographicalMapping/datasets/completeLRTB-IntegratedSct.rds")
saveRDS(all.combined.sct, "/Users/josongco/Desktop/topographicalMapping/datasets/completeLR-IntegratedSct.rds")
saveRDS(all.combined.sct, "/Users/josongco/Desktop/topographicalMapping/datasets/completeTB-IntegratedSct.rds")
all.combined.sct <- readRDS("/Users/josongco/Desktop/topographicalMapping/datasets/completeLRTB-IntegratedSct.rds")
```


# Neurons UMAP from Songco-Casey et al., 2022
```{r}
neurons_ordered <- readRDS("/Users/josongco/Desktop/Cell Types Seurat Code & Files/FinalNeuronsUMAP-ordered.rds") 
DimPlot(neurons_ordered, label = TRUE, cols = UMAPcolors)

neurons_ordered_renamed <- RenameIdents(neurons_ordered, '18' = "immature", '19' = "immature", '20' = "immature", '27' = "immature", '22' = "immature", '23' = "immature", '24' = "immature", '25' = "immature", '26' = "immature", 
                                         '10' = "dopaminergic", '11' = "dopaminergic", '12' = "dopaminergic", '13' = "dopaminergic", '14' = "dopaminergic", '15' = "dopaminergic", '16' = "dopaminergic", '17' = "dopaminergic",
                                         '5' = "dopaminergic+glutamatergic", '6' = "dopaminergic+glutamatergic", '36' = "dopaminergic+glutamatergic",
                                         '28' = "glutamatergic", '29' = "glutamatergic", '30' = "glutamatergic", '31' = "glutamatergic", '32' = "glutamatergic", '33' = "glutamatergic", '34' = "glutamatergic",
                                         '7' = "cholinergic", '8' = "cholinergic", '9' = "cholinergic", '35' = "cholinergic",
                                         '4' = "octopaminergic", 
                                         '37' = "orcokinin")
UMAPcolors_lumped <- c('hotpink', 'steelblue', 'darkseagreen', 'goldenrod', 'sienna1', 'red', 'purple4')
DimPlot(neurons_ordered_renamed, label = TRUE, cols = UMAPcolors_lumped)

levels(neurons_ordered_renamed)
neurons_ordered_renamed$seurat_clusters <- levels(neurons_ordered_renamed)
```

```{r}
#all.combined.sct_trialNames <- all.combined.sct
#all.combined.sct_trialNames$seurat_clusters <- levels(neurons_ordered_renamed)

#p1 <- DimPlot(all.combined.sct_trialNames, reduction = "umap", group.by = "orig.ident")
#p2 <- DimPlot(all.combined.sct_trialNames, reduction = "umap", group.by = "seurat_clusters", label = TRUE, repel = TRUE)
#p1 + p2

p1 <- DimPlot(neurons_ordered_renamed, group.by = "orig.ident", label = TRUE)
p2 <- DimPlot(neurons_ordered_renamed, group.by = "seurat_clusters", label = TRUE, cols = UMAPcolors_lumped) # doesn't reflect what I was hoping it would
p1 + p2

saveRDS(neurons_ordered_renamed, "/Users/josongco/Desktop/topographicalMapping/datasets/neuronsLumped.rds")
```



```{r}
DefaultAssay(all.combined) <- "SCT"
pdf(file = "/Users/josongco/Desktop/topographicalMapping/sct_LR.top30clusters.pdf",
    width = 10, # The width of the plot in inches
    height = 10) # 
for (m in 0:(nlevels(all.combined$seurat_clusters)-1)){
  clustermarkers_i <- FindMarkers(all.combined, ident.1 = m, logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
  clustermarkers_i$diff_exp <- (clustermarkers_i$pct.1/clustermarkers_i$pct.2)
  
  clustermarkers_i <- clustermarkers_i[order(-clustermarkers_i$diff_exp), ]
top30_genes <- rownames(clustermarkers_i)[1:30]
top30_genes <- data.frame(top30_genes)
  top30_genes<-addGeneNamesTable(top30_genes,top30_genes$top30_genes,gene_master$obgene, gene_master$merge_desc)
  
  print(top30_genes) #top30
  p<-DotPlot(all.combined,features=top30_genes$top30_genes, cols = c("blue","red"), split.by = "orig.ident") + RotatedAxis() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7)) + NoLegend() #top30
  p<-replaceLabelsX(p,gene_master$obgene,gene_master$merge_desc)
  p <- p+ coord_flip()
  p <- p+ ggtitle(paste('top30 for cluster', m, sep = " ")) 
  print(p)  
}

dev.off()
```



```{r}
DefaultAssay(allcombined.celltypesample) <- "SCT"
pdf(file = "/Users/josongco/Desktop/topographicalMapping/sct_TB.top50samples-avgLog2FC.pdf",
    width = 10, # The width of the plot in inches
    height = 10) # 
for (m in 0:(nlevels(allcombined.celltypesample$seurat_clusters)-1)){
  clustermarkers_i <- FindMarkers(allcombined.celltypesample, ident.1 = paste(m, "Sample3", sep = "_"), ident.2 = paste(m, "Sample4", sep = "_"), logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
  clustermarkers_i$diff_exp <- (clustermarkers_i$pct.1/clustermarkers_i$pct.2)
  
  clustermarkers_i <- clustermarkers_i[order(-(abs(clustermarkers_i$avg_log2FC))), ]
top50_genes <- rownames(clustermarkers_i)[1:50]
top50_genes <- data.frame(top50_genes)
  top50_genes<-addGeneNamesTable(top50_genes,top50_genes$top50_genes,gene_master$obgene, gene_master$merge_desc)
  
  #print(top30_genes) #top30
  p<-DotPlot(allcombined.celltypesample.order,features=top50_genes$top50_genes, cols = c("blue","red"), split.by = "orig.ident") + RotatedAxis() + theme(axis.text.x = element_text(size = 6)) + theme(axis.text.y = element_text(size = 6)) + theme(legend.title = element_text(size = 7)) + theme(legend.text = element_text(size = 7)) + NoLegend() #top30
  p<-replaceLabelsX(p,gene_master$obgene,gene_master$merge_desc)
  p <- p+ coord_flip()
  #p <- p+ ggtitle(paste('topvbot top50 for ident', m, sep = " ")) 
  print(p)  
}

dev.off()


FeaturePlot(allcombined.celltypesample.order, features = top50_genes$top50_genes[1:5])


allcombined.celltypesample.order <- allcombined.celltypesample
levels(allcombined.celltypesample.order) <- order


example.response <- FindMarkers(allcombined.celltypesample, ident.1 = "0_Sample3", ident.2 = "0_Sample4", verbose = FALSE)
head(example.response, n = 15)
example.response <- rownames(example.response)
example.response <- data.frame(example.response)
example.response$gene <- example.response$example.response
example.response <- addGeneNamesTable(example.response, example.response$gene, example.response$gene, gene_master$merge_desc)
top30_btwn <- head(example.response, n = 30)
```

